NumLin: Linear Types for Linear Algebra

### 2018-11-05 write-up/paper/implementation.tex
Add interpretation and proof
Start on Soundness Theorem

    \item \textbf{Matrix Expressions} are also desugared into the abstract
        syntax tree through some simple pattern-matching.

@@

    \item \textbf{Code Generation}. The abstract syntax tree is translated into
        OCaml, with a few `optimisations' to produce more readable code. This
        process is type-preserving: \lang's type system is embedded into
        OCaml's (Figure~\ref{fig:type_grammar}), and so the OCaml type checker
        acts as a sanity check on the generated code.

@@

Desugaring is conventional, outlined in Figure~\ref{fig:lang_desugar}.  Matrix
expression are translated into BLAS/ LAPACK calls via purely syntactic
pattern-matching, outlined in Figure~\ref{fig:lang_matexp}.

\subsubsection{Type checking} is mostly standard for a linearly typed language,
with the exception of fractional permission inference. Because all functions
must have their argument types explicitly annotated, inferring the correct
fraction at a call-site is simply a matter of unification.

@@

\textsc{What more can/should we say here?}

@@

\begin{figure}[t]
\begin{center}
    \begin{tabular}{rcl}
        \alsocell[b]{l}{\highl{let Many x = x in} \\
        \highl{let Many x = Many (Many x) in <exp>} } &
        $\Rightarrow$ &
        \highl{<exp>}
    \\ \\
        \alsocell[c]{l}{\highl{let Many x = <exp> in} \\
            \highl{let Many x = Many (Many x) in <body>} \\
            \textbf{OR} \highl{let Many x = Many <exp> in <body>} \\
            \textbf{OR} \highl{(fun x : t -> <body>) <exp>}} &
        $\Rightarrow$ &
        \highl{let x = <exp> in <body>}
    \\ \\
        \alsocell[c]{l}{\highl{(* fixp = fix (f, x:t, <exp> : t') *)} \\
            \highl{let Many f = Many fixp in <body>} \\
            \textbf{OR} \highl{let f = fixp in <body>} } &
        $\Rightarrow$ &
        \highl{let rec f x = <exp> in <body>}
    \end{tabular}
\end{center}
\caption{Removing redundant re-bindings during tranlation to OCaml.}\label{fig:lang_optimise}
\end{figure}

>>>>

    \item \textbf{Matrix Expressions} are also desugared into the abstract
        syntax tree through pattern-matching.

@@

    \item \textbf{Code Generation}. The abstract syntax tree is translated into
        OCaml, with a few `optimisations' to produce more readable code. This
        process is type-preserving: \lang's type system is embedded into
        OCaml's (Figure~\ref{fig:type_grammar}), so the OCaml type checker
        acts as a sanity check on the generated code.

@@

Desugaring is conventional, outlined in Appendix~\ref{sec:lang_desugar}.  Matrix
expression are translated into BLAS/ LAPACK calls via purely syntactic
pattern-matching, outlined in Figure~\ref{fig:lang_matexp}.

\subsubsection{Type checking} is mostly standard for a linearly typed language,
with the exception of fractional permission inference. By restricting fractions
to be non-positive integer powers of two, we only need to keep track of the
logarithm of the fractions used. Explicit sharing and unsharing removes the
need for performing dataflow analysis. As a result, all fractional arithmetic
can be solved with unification, and in doing so, fractions become directly
usable in \lang's type-system as opposed to a convenient theoretical tool.

Because all functions must have their argument types explicitly annotated,
inferring the correct fraction at a call-site is simply a matter of
unification. We strongly suspect that \emph{full-inference of fractional
permissions is similarly just matter of unification} (thanks to an
experimental implementation of just this feature), even though the formal
system we present here is for an explicitly-typed language.

@@

@@

\begin{figure}[t]
\begin{center}
    \begin{tabular}{rcl}
        \alsocell[b]{l}{\highl{let Many x = x in} \\
        \highl{let Many x = Many (Many x) in <exp>} } &
        $\Rightarrow$ &
        \highl{<exp>}
    \\ \\
        \alsocell[c]{l}{\highl{let Many x = <exp> in} \\
            \highl{let Many x = Many (Many x) in <body>} \\
            \textbf{OR} \highl{let Many x = Many <exp> in <body>} \\
            \textbf{OR} \highl{(fun x : t -> <body>) <exp>}} &
        $\Rightarrow$ &
        \highl{let x = <exp> in <body>}
    \\ \\
        \alsocell[c]{l}{\highl{(* fixp = fix (f, x:t, <exp> : t') *)} \\
            \highl{let Many f = Many fixp in <body>} \\
            \textbf{OR} \highl{let f = fixp in <body>} } &
        $\Rightarrow$ &
        \highl{let rec f x = <exp> in <body>}
    \end{tabular}
\end{center}
\caption{Removing redundant re-bindings during translation to OCaml.}\label{fig:lang_optimise}
\end{figure}
